---
title: "Determinants of alcohol use during pregnancy"
author: "Sylvia Roozen & Gjalt-Jorn Peters"
date: "`r format(Sys.time(), '%H:%M:%S on %Y-%m-%d %Z (GMT%z)')`"
output: html_document
---

```{r setup, include=FALSE}

########################################################################
### Load packages
########################################################################

require('userfriendlyscience');  ### For convenience functions, e.g. 'safeRequire'
safeRequire('here');             ### To easily access files using 'relative paths'
safeRequire('plyr');             ### For easily processing and restructuring data
safeRequire('dplyr');            ### For case_when
safeRequire('purrr');            ### For quickly getting value from list of lists with map
safeRequire('magrittr');         ### For the pipe operator (%>%)
safeRequire('ggplot2');          ### For plots
safeRequire('pander');           ### Pretty output in Markdown
safeRequire('knitr');            ### Tables in Markdown
safeRequire('data.tree');        ### For hierarchical data
safeRequire('devtools');         ### To install metabefor from github repo
                                 ### ... Which we then do here:
devtools::install_github("Matherion/metabefor");
require('metabefor');

########################################################################
### Settings
########################################################################

### By default hide R code
knitr::opts_chunk$set(echo = FALSE);

### Set path for query hit exports
queryHitExportPath <- here::here("10-queries");

### Set path for screening
screeningPath <- here::here("20-screening");

### Set path for extraction script template
extractionPath <- here::here("40-extraction-scripts", "final-extractions-quantitative");

### Set working path
workingPath <- here::here("60-intermediate-results");

########################################################################
### Valid values to be specified for extracted entities
########################################################################

allowedValues.variable = c("variable",
                           "moment",
                           "type",
                           "datatype",
                           "values",
                           "labels",
                           "psytype",
                           "dependent",
                           "direction",
                           "parent",
                           "description",
                           "comment",
                           'dataCollectionMethod',
                           'timeframe',
                           'period',
                           'determinant_original',
                           'determinant_recode',
                           'determinant_target',
                           'determinant_direction',
                           'determinant_consistent',
                           'determinant_inconsistent');

```

# Processing the extraction scripts

## Initial parsing {.tabset}

### Overview

Click the 'Logs' tab to see the detailed logs.

```{r process-extraction-scripts, warning=FALSE}

###########################################################################
### Start processing
###########################################################################

### Process all extracted articles
processExtractedFiles_log <-
  capture.output(extractedData <-
                   processExtractedFiles(extractionPath,
                                         allowedValues.variable=allowedValues.variable)) %>%
  paste(collapse="\n");
  
### Store in a file
writeLines(processExtractedFiles_log,
           file.path(workingPath, "processing-extraction-scripts-log.txt"));

### View numbers of qualitative and quantitative studies -
### there should be no qualitative studies in this dataset
pander(freq(unlist(lapply(extractedData$intermediate$processedObjects, function(x) return(x$datacollectionmethod)))));

### View which extraction scripts still miss the method
which(is.na(lapply(extractedData$intermediate$processedObjects, function(x) return(x$datacollectionmethod))));

```

### Logs

```{r detailed-parsing-logs}

### Show the output
cat(processExtractedFiles_log);

```

### Determinant categorization

## Second level categorization

```{r second-level-categorization}

l_ply(seq_along(extractedData$intermediate$processedObjects), function(studyIndex) {
  return(l_ply(seq_along(extractedData$intermediate$processedObjects[[studyIndex]]$variable), function(varIndex) {

    originalCoding <- extractedData$intermediate$processedObjects[[studyIndex]]$variable[[varIndex]]$determinant_recode;
    originalName <- extractedData$intermediate$processedObjects[[studyIndex]]$variable[[varIndex]]$variable;
    
    #print(grepl('drinking', originalName, ignore.case = TRUE));

    extractedData$intermediate$processedObjects[[studyIndex]]$variable[[varIndex]]$determinant_recode_2 <<-
      case_when(grepl('attitud', originalCoding, ignore.case = TRUE) ~
                  "attitude",
                grepl('moral norm', originalCoding, ignore.case = TRUE) ~
                  "moralnorm",
                grepl('personal norm', originalCoding, ignore.case = TRUE) ~
                  "personalnorm",
                grepl('descriptive norm', originalCoding, ignore.case = TRUE) ~
                  "descriptivenorm",
                grepl('norm', originalCoding, ignore.case = TRUE) ~
                  "injunctivenorm",
                grepl('denial', originalCoding, ignore.case = TRUE) ~
                  "denial",
                grepl('expectation', originalCoding, ignore.case = TRUE) ~
                  "attitude",
                grepl('social support', originalCoding, ignore.case = TRUE) ~
                  "socialsupport",
                grepl('guilt', originalCoding, ignore.case = TRUE) ~
                  "guilt",
                grepl('intention', originalCoding, ignore.case = TRUE) ~
                  "intention",
                grepl('knowledge', originalCoding, ignore.case = TRUE) ~
                  "knowledge",
                grepl('motivation to comply', originalCoding, ignore.case = TRUE) ~
                  "motivationtocomply",
                grepl('response efficacy', originalCoding, ignore.case = TRUE) ~
                  "responseefficacy",
                grepl('risk perception', originalCoding, ignore.case = TRUE) ~
                  "riskperception",
                grepl('self.?efficacy', originalCoding, ignore.case = TRUE) ~
                  "selfefficacy",
                grepl('severity', originalCoding, ignore.case = TRUE) ~
                  "severity",
                grepl('susceptibility', originalCoding, ignore.case = TRUE) ~
                  "susceptibility",
                grepl('drinking', originalName, ignore.case = TRUE) ~
                  "behavior",
                grepl('alcohol', originalName, ignore.case = TRUE) ~
                  "behavior",
                is.null(originalCoding) ~
                  "NULL",
                is.na(originalCoding) ~
                  "NA",
                TRUE ~
                  "NO MATCH!");

  }));
});

prettyLabels <-
  c(attitude = "Attitude",
    moralnorm = "Moral Norm",
    personalnorm = "Personal Norm",
    descriptivenorm = "Descriptive Norm",
    injunctivenorm = "Injunctive Norm",
    subjectivenorm = "Subjective Norm",
    perceivednorm = "Perceived Norm",
    denial = "Denial",
    socialsupport = "Social Support",
    guilt = "Guilt",
    intention = "Intention",
    knowledge = "Knowledge",
    motivationtocomply = "Motivation To Comply",
    responseefficacy = "Response Efficacy",
    eppmefficacy = 'EPPM`s efficacy',
    outcomeexpectation = "Outcome Expectation",
    riskperception = "Risk Perception",
    outcomeexpectations = "Outcome Expectation",
    selfefficacy = "Self-Efficacy",
    severity = "Severity",
    intention = "Intention",
    agency = "Agency",
    susceptibility = "Susceptibility",
    behavior = "Alcohol Use");

```

```{r adding-values-and-labels-to-univariate-and-associations}

l_ply(seq_along(extractedData$intermediate$processedObjects), function(studyIndex) {
  return(l_ply(seq_along(extractedData$intermediate$processedObjects[[studyIndex]]$univariate), function(resultIndex) {

    referredVariableName <- extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]]$variable;
    
    referredVariableExtraction <-
      extractedData$intermediate$processedObjects[[studyIndex]]$variable[[
        which(unlist(purrr::map(extractedData$intermediate$processedObjects[[studyIndex]]$variable, 'variable')) == referredVariableName)
      ]];
    
    referredValueLabelVector <-
      referredVariableExtraction$labels;

    names(referredValueLabelVector) <-
      as.character(referredVariableExtraction$values);

    extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]]$valueLabel <<-
      unname(referredValueLabelVector[as.character(extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]]$value)]);
    
    extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]]$description <<-
      referredVariableExtraction$description;

    extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]]$determinantId <<-
      referredVariableExtraction$determinant_recode_2;
    
    extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]]$study <<-
      names(extractedData$intermediate$processedObjects)[studyIndex];
    
  }));
});

l_ply(seq_along(extractedData$intermediate$processedObjects), function(studyIndex) {
  return(l_ply(seq_along(extractedData$intermediate$processedObjects[[studyIndex]]$association), function(resultIndex) {

    referredVariableName <- extractedData$intermediate$processedObjects[[studyIndex]]$association[[resultIndex]]$variable1;

    referredVariableExtraction <-
      extractedData$intermediate$processedObjects[[studyIndex]]$variable[[
        which(unlist(purrr::map(extractedData$intermediate$processedObjects[[studyIndex]]$variable, 'variable')) == referredVariableName)
      ]];
    
    referredValueLabelVector <-
      referredVariableExtraction$labels;

    names(referredValueLabelVector) <-
      as.character(referredVariableExtraction$values);

    extractedData$intermediate$processedObjects[[studyIndex]]$association[[resultIndex]]$description <<-
      referredVariableExtraction$description;

    extractedData$intermediate$processedObjects[[studyIndex]]$association[[resultIndex]]$determinantId <<-
      referredVariableExtraction$determinant_recode_2;
    
    extractedData$intermediate$processedObjects[[studyIndex]]$association[[resultIndex]]$study <<-
      names(extractedData$intermediate$processedObjects)[studyIndex];
    
  }));
});

```

## Overview of included operationalisations and how they were recoded {.tabset}

```{r operationalisation-recoding}

af <- function(value) return(ifelse(is.null(value) ||
                                    is.na(value), "", value));

recodedOperationalisations <-
  ldply(extractedData$intermediate$processedObjects, function(study) {
    return(ldply(study$variable, function(x) {
      return(data.frame(varName = x$variable,
                        varDet_original = af(x$determinant_original),
                        varDet_recoded = af(x$determinant_recode),
                        varDet_recoded2 = af(x$determinant_recode_2),
                        varDet_targetObject = af(x$determinant_target),
                        varDet_direction = af(x$determinant_direction),
                        varDet_targetBehavior = af(x$determinant_consistent),
                        varDet_contrastBehavior = af(x$determinant_inconsistent),
                        varDescription = af(x$description),
                        stringsAsFactors=FALSE));
    }));
  }, .id="study");

### Set all missing (NA) recodes to 'behavior'
recodedOperationalisations$varDet_recoded2 <-
  ifelse(is.na(recodedOperationalisations$varDet_recoded2),
         'behavior',
         recodedOperationalisations$varDet_recoded2);

write.csv(recodedOperationalisations,
          file.path(workingPath, "recoded-operationalisations.csv"),
          row.names=FALSE);

```

### Read determinant definitions and hierarchy

```{r determinant-hierarchy}

### Import recoding specifications
determinantHierarchy.specification <-
  read.csv(here::here('55-recoding', 'recoding.csv'));

### Get subset with only children and parents
detH <- determinantHierarchy.specification[, c('identifier', 'parents')];

### Remove all but first parent where multiple are listed
detH$parents <- gsub(', .*', '', detH$parents);

### Convert to network
determinantHierarchy <- FromDataFrameNetwork(detH);

### Count operationalisations
operCount <- table(recodedOperationalisations$varDet_recoded2);

### Add this count to the determinant hierarchy
l_ply(names(operCount),
      function(determinantIdentifier) {
        Do(list(FindNode(determinantHierarchy,
                         determinantIdentifier)),
           function(nd) {
             nd$nrOfItems <- operCount[determinantIdentifier];
           });
  });

### Set number of operationalisations to 0 where it's NA
determinantHierarchy$Do(function(nd) {
  nd$Set(nrOfItems =
    ifelse(is.na(nd$Get('nrOfItems')),
           0,
           nd$Get('nrOfItems')));
});

### Add items accumulated over children;
### First create a custom recursive function (Aggregate
### doesn't seem to be able to cut it for some reason)
getTotalNrOfItems <- function(node) {
  if (node$isLeaf) {
    return(node$nrOfItems);
  } else {
    return(sum(node$nrOfItems,
               unlist(lapply(node$children,
                             getTotalNrOfItems))));
  }
}

### Apply this function
determinantHierarchy$Do(function(nd) {
  nd$accumulatedItems <- getTotalNrOfItems(nd);
  }, traversal = "post-order");

#print(determinantHierarchy, 'nrOfItems', 'accumulatedItems');

### Remove ('prune') unused determinants from the tree
invisible(Prune(determinantHierarchy,
      function(nd) {
        if (nd$accumulatedItems == 0) {
          return(FALSE);
        } else {
          return(TRUE);
        }
      }));

#print(determinantHierarchy, 'nrOfItems', 'accumulatedItems');

### Set labels as both name and number of items
SetNodeStyle(determinantHierarchy,
             shape = "box",
             style = "rounded",
             fontname = "Arial",
             label = function(x) paste0(prettyLabels[x$name], "\n",
                                        "(",
                                        x$nrOfItems,
                                        " items; ",
                                        x$accumulatedItems,
                                        " incl. subdeterminants)"));

SetGraphStyle(determinantHierarchy,
              rankdir = "RL");

SetEdgeStyle(determinantHierarchy,
             color = "#000000",
             style="solid",
             dir="back",
             penwidth = 2);

### Plot hierarchy
plot(determinantHierarchy);

DiagrammeR::export_graph(data.tree::ToDiagrammeRGraph(determinantHierarchy),
                         file_name = file.path(workingPath,
                                               "operationalisations-overview.svg"),
                         file_type = "svg");

```


```{r add-results-to-determinant-hierarchy}

l_ply(seq_along(extractedData$intermediate$processedObjects), function(studyIndex) {
  return(l_ply(seq_along(extractedData$intermediate$processedObjects[[studyIndex]]$univariate), function(resultIndex) {

    univarExtraction <- extractedData$intermediate$processedObjects[[studyIndex]]$univariate[[resultIndex]];
    
    determinantNode <- FindNode(determinantHierarchy,
                                univarExtraction$determinantId);
    
    if (is.null(determinantNode$univar)) {
      determinantNode$univar <- as.data.frame(univarExtraction);
    } else {
      determinantNode$univar <- plyr::rbind.fill(determinantNode$univar,
                                                 as.data.frame(univarExtraction));
    }
    
    determinantNode$univar <-
      Filter(function(x) !all(is.na(x)),
             determinantNode$univar);

  }));
});

l_ply(seq_along(extractedData$intermediate$processedObjects), function(studyIndex) {
  return(l_ply(seq_along(extractedData$intermediate$processedObjects[[studyIndex]]$association), function(resultIndex) {

    assocExtraction <- extractedData$intermediate$processedObjects[[studyIndex]]$association[[resultIndex]];
    
    print(assocExtraction$study);
    print(assocExtraction$variable1);
    print(assocExtraction$variable2);
    print(assocExtraction$determinantId);
    
    determinantNode <- FindNode(determinantHierarchy,
                                assocExtraction$determinantId);
    
    if (is.null(determinantNode$assoc)) {
      determinantNode$assoc <- as.data.frame(assocExtraction);
    } else {
      determinantNode$assoc <- plyr::rbind.fill(determinantNode$assoc,
                                                 as.data.frame(assocExtraction));
    }
    
    determinantNode$assoc <-
      Filter(function(x) !all(is.na(x)),
             determinantNode$assoc);

  }));
});

```

```{r determinant-types, results='asis'}

### Add operationalisations and metadata about the
### operationalisations
invisible(determinantHierarchy$Do(function(nd) {
  
  nd$targetObjects_other <-
    sum(grepl('other',
        recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                         'varDet_targetObject']));
  nd$targetObjects_self <-
    sum(grepl('self',
        recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                         'varDet_targetObject']));
  nd$targetObjects_unclear <-
    sum(grepl('unclear',
        recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                         'varDet_targetObject']));
  nd$targetObjects_unspecified <-
    sum(grepl('unspecified',
        recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                         'varDet_targetObject']));
  nd$targetObjects_neutral <-
    sum(grepl('neutral',
        recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                         'varDet_targetObject']));

  nd$targetObjects <-
    table(recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                     'varDet_targetObject']);

  nd$directions_negative <-
    sum(grepl('negative',
        recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_direction']));
  nd$directions_negative <-
      sum(grepl('negative',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_direction']));
  
  nd$directions_neutral <-
      sum(grepl('neutral',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_direction']));
  
  nd$directions_positive <-
      sum(grepl('positive',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_direction']));

  nd$directions <-
    table(recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                     'varDet_direction']);

  nd$targetBehaviors_A <-
    sum(grepl('A',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_targetBehavior']));

  nd$targetBehaviors_M <-
    sum(grepl('M',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_targetBehavior']));

  nd$targetBehaviors_D <-
    sum(grepl('D',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_targetBehavior']));

  nd$targetBehaviors <-
    table(recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                     'varDet_targetBehavior']);

  nd$contrastBehaviors_A <-
    sum(grepl('A',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_contrastBehavior']));

  nd$contrastBehaviors_M <-
    sum(grepl('M',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_contrastBehavior']));

  nd$contrastBehaviors_D <-
    sum(grepl('D',
          recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                           'varDet_contrastBehavior']));

  nd$contrastBehaviors <-
    table(recodedOperationalisations[recodedOperationalisations$varDet_recoded2 == nd$name,
                                     'varDet_contrastBehavior']);
  
}));

kablePrintNode <- function(node, ...) {
  null <- capture.output(res <- invisible(kable(print(node, ...))));
  return(res);
}

kablePrintNode(determinantHierarchy,
            'targetObjects_self',
            'targetObjects_other',
            'targetObjects_unclear',
            'targetObjects_unspecified',
            'targetObjects_neutral');

kablePrintNode(determinantHierarchy,
            'directions_negative',
            'directions_neutral',
            'directions_positive');

kablePrintNode(determinantHierarchy,
            'targetBehaviors_A',
            'targetBehaviors_M',
            'targetBehaviors_D');

kablePrintNode(determinantHierarchy,
              'contrastBehaviors_A',
              'contrastBehaviors_M',
              'contrastBehaviors_D');

#print(determinantHierarchy,'operationalisations');


```

### Overview

Click the tabs above to view the operationalisations.

```{r number-of-operationalisations}

kable(table(recodedOperationalisations[order(recodedOperationalisations$varDet_recoded), ]$varDet_recoded));

#recodedOperationalisations[, c('varName', 'study', 'varDet_original', 'varDet_recoded', 'varDet_recoded2')];

```

### Study order

```{r recodedOperationalisations-study-order}

kable(recodedOperationalisations);

```

### Ordered by recoded determinant

```{r recodedOperationalisations-recodedOperationalisation-order}

kable(recodedOperationalisations[order(recodedOperationalisations$varDet_recoded), ]);

```

## Number of extracted univariate and bivariate results

```{r result-counting}

univarCount <-
  lapply(extractedData$intermediate$processedObjects, function(x)
    return(ifelse(is.null(x$univariate) || is.na(x$univariate), 0, length(x$univariate))));

univarCount <- cbind(names(univarCount), univarCount);
row.names(univarCount) <- NULL;
kable(univarCount);

cat("\n\nAvailable bivariate results:",
     vecTxtQ(sort(unique(unlist(lapply(extractedData$intermediate$processedObjects,
                                       function(x)
    return(ifelse(is.null(x$univariate) ||
                  is.na(x$univariate) ||
                  length(is.null(x$univariate))==0,
                  0,
                  list(lapply(x$univariate, function(y) {
                    if (sum(!is.na(y)) > 0) {
                      return(names(y)[!is.na(y)]);
                    } else {
                      return(NULL);
                    }
                  }))))))))));

assocCount <-
  lapply(extractedData$intermediate$processedObjects, function(x)
    return(ifelse(is.null(x$association) || is.na(x$association), 0, length(x$association))));

assocCount <- cbind(names(assocCount), assocCount);
row.names(assocCount) <- NULL;
kable(assocCount);

cat("\n\nAvailable associations:",
     vecTxtQ(sort(unique(unlist(lapply(extractedData$intermediate$processedObjects,
                                       function(x)
    return(ifelse(is.null(x$association) ||
                  is.na(x$association) ||
                  length(is.null(x$association))==0,
                  0,
                  list(lapply(x$association, function(y) {
                    if (sum(!is.na(y)) > 0) {
                      return(names(y)[!is.na(y)]);
                    } else {
                      return(NULL);
                    }
                  }))))))))));

```

## Detailed univariate results

```{r detailed-univariate-results}

### Add sample information to univariate data
sampleInfo <-
  names(extractedData$intermediate$processedObjects[[1]]$sample);
extractedData$output$univariate <-
  addSampleInfo(listOfObjects = extractedData$intermediate$processedObjects,
                dataframe = extractedData$output$univariate,
                variableToAdd = setdiff(sampleInfo,
                                        names(extractedData$output$univariate)));

### Add variable (operationalisation) information to univariate data
extractedData$output$univariate <-
  addVariableInfo(variablesDataframe = extractedData$output$variable,
                  dataframe = extractedData$output$univariate,
                  variableToAdd = setdiff(names(extractedData$output$variable),
                                          names(extractedData$output$univariate)));

extractedData$output$univariate$percentage <-
  round(extractedData$output$univariate$percentage,
        2);

knitr::kable(extractedData$output$univariate[, c('study',
                                                 'variable',
                                                 'determinant_recode',
                                                 'percentage',
                                                 'determinant_target',
                                                 'determinant_direction',
                                                 'determinant_consistent',
                                                 'determinant_inconsistent')]);

# Percentages -> vector
# Labels -> strsplit

```

## Detailed association results

```{r detailed-association-results}

### Add sample information to association data
sampleInfo <-
  names(extractedData$intermediate$processedObjects[[1]]$sample);
extractedData$output$association <-
  addSampleInfo(listOfObjects = extractedData$intermediate$processedObjects,
                dataframe = extractedData$output$association,
                variableToAdd = setdiff(sampleInfo,
                                        names(extractedData$output$association)));

### Add variable (operationalisation) information to association data
extractedData$output$association <-
  addVariableInfo(variablesDataframe = extractedData$output$variable,
                  dataframe = extractedData$output$association,
                  variableToAdd = setdiff(names(extractedData$output$variable),
                                          names(extractedData$output$association)));

knitr::kable(extractedData$output$association[, c('study',
                                                  'variable1',
                                                  'variable2')]);

```

# Results organised by determinant structure

```{r results-by-determinant-structure, results='asis'}

invisible(determinantHierarchy$Do(function(node) {
  cat0("\n\n## ", prettyLabels[node$name], "\n\n");

  cat0("\n\n### Univariate results\n\n");
  
  if (is.data.frame(node$univar)) {
    print(knitr::kable(node$univar));
  } else {
    cat0("\n\nNo univariate results.\n\n");
  }
  
  cat0("\n\n### Associations\n\n");

  if (is.data.frame(node$assoc)) {
    print(knitr::kable(node$assoc));
  } else {
    cat0("\n\nNo associations.\n\n");
  }
  
}, traversal = 'pre-order'));

```

